---
title: "Consumer vs Producer Egg Prices Over Time"
author: "Chad Gueli, Alina Martinez, Amy Philip, and Xuliejun Ren"
date: "12/6/2021"
output:
    beamer_presentation:
        theme: "AnnArbor"
        colortheme: "dolphin"
        fonttheme: "structurebold"
---

```{r setup, include=FALSE}
library(MTS)
library(seastests)
library(vars)
library(tseries)
library(tidyverse)
library(feather)

data <- read_feather("C:/Users/Gloria/Downloads/orig_data.feather") %>%
    select(index, c_eggs, p_eggs) %>%
    mutate(ceggs=c_eggs, peggs=p_eggs/100, .keep="unused")

pts <- ts(data$peggs, frequency=12, start=c(2000,1))
cts <- ts(data$ceggs, frequency=12, start=c(2000,1))
eggts <- cbind(pts, cts)

pdiff <- diff(pts, lag=12)
cdiff <- diff(cts, lag=12)

diffed <- cbind(pdiff, cdiff)

knitr::opts_chunk$set(echo = FALSE)
```

##  PPI of Eggs Over Time
```{r peg}
plot.zoo(pts, col=4, las=1, main="Producer Price of Eggs", ylab="Price ($)")
```

##  CPI of Eggs Over Time
```{r ceg}
plot.zoo(cts, col=4, las=1, main="Consumer Price of Eggs", ylab="Price ($)")
```

## Auto-Correlation of Producer Egg Prices
```{r ppppp}
acf(pts)
pacf(pts)
```

## Auto-Correlation of Consumer Egg Prices
```{r cccccc}
acf(cts)
pacf(cts)
```

## Kruskal-Wallis Test

$H_0$: No seasonality (i.e. medians are the same at different lags)

- Producer Seasonality
    - Reject at the 0.01 level
    
\tiny
```{r}
kw(pts)
```

\normalsize
- Consumer Seasonality
    - Reject at the 0.01 level

\tiny
```{r}
kw(cts)
```

## Remove Seasonality

We difference with a lag of 12, to remove seasonality.

- Producer Seasonality
    - Fail to reject at the 0.01 level
    
\tiny
```{r}
kw(pdiff)
```

\normalsize
- Consumer Seasonality
    - Fail to reject at the 0.01 level

\tiny
```{r}
kw(cdiff)
```

\normalsize
- After differencing, we are unable to conclude that there is seasonality.

## Plot of 12-Month Difference in Producer Price
```{r, out.width="80%"}
plot(pdiff, ylab="12-Month Price Difference")
```

- Series looks to be centered at 0, suggesting stationarity.

## Plot of 12-Month Difference in Consumer Price

```{r, out.width="80%"}
plot(cdiff, ylab="12-Month Price Difference")
```

- Process appears to have mean 0, implying stationarity.

## Augmented Dickey-Fuller Test

$H_0$: Time series is non-stationary.

- Producer Price
    - Reject at least at the 0.01 level.
    
\tiny
```{r, warning=FALSE}
adf.test(pdiff)
```
\normalsize
- Consumer Price
    - Reject at least at the 0.01 level.

\tiny
```{r, warning=FALSE}
adf.test(cdiff)
```

## Vector Auto-Regressive (VAR) Model
- For multivariate series
- Relation of lags and lags of other variables
- Before we need to check if one causes the other

## Granger Test

$H_0$: No sequence causes another.

\tiny
```{r granger}
GrangerTest(diffed, p=5)
```

\normalsize
- Reject null hypothesis at 0.01 level and concluded there is causality between the time series.

## Selecting Lag

```{r}
VARselect(diffed)$selection
```
We use the Schwarz Criterion or BIC, and will fit an VAR process of order 5.

## VAR model


```{r  fit model, echo=TRUE}
vm <- vars::VAR(diffed, p=5)
```
We fit the model with the `VAR` function from the vars package.

## Egg Price Coefficients

- Consumer

\tiny
```{r}
vm$varresult$cdiff
```

\normalsize
- Producer

\tiny
```{r}
vm$varresult$pdiff
```

## Jarque-Bera Normality Tests

$H_0$: residual skewness and kurtosis are consistent with normality

- Skewness
    - Fail to reject assumption that residual skewness is consistent with normality.

\tiny
```{r norm skew}
ntest <- normality.test(vm)
ntest$jb.mul$Skewness
```

\normalsize
- Kurtosis
    - Reject assumption that residual kurtosis is consistent with normality.

\tiny
```{r norm kurt}
ntest$jb.mul$Kurtosis
```


## Test of Residual Serial Correlation

$H_0$: No residual serial correlation

- Breusch-Goldfrey Test
    - Score test

\tiny
```{r BG}
serial.test(vm, type="BG")
```

\normalsize
- Edgerton-Shukur Test
    - Likelihood-Ratio (F) test

\tiny
```{r ES}
serial.test(vm, type="ES")
```

\normalsize
- Both tests reject assumption at the 0.01 level

## Discussion of Problems
- The residuals are both serially correlated and non-normal, these are likely symptoms of the same problem

- There seems to exist heteroscedasticity in the residuals, preventing an analytic comparison of the series.


## Stability Plot for De-Seasoned Producer Egg Price
```{r stability}
stab <- stability(vm)
plot(stab$stability$pdiff)
```

## Stability Plot for De-Seasoned Consumer Egg Price
```{r stability2}
plot(stab$stability$cdiff)
```

## Our Model

```{r mod plot, out.width="45%", out.height="100%"}
plot(vm)
```


## The Fan Chart
```{r fan chart, out.height="80%"}
var.f10 <- predict(vm, n.ahead = 10, ci = 0.99)
fanchart(var.f10, main=c("PPI Eggs", "CPI Eggs"))
```

